---
title: "Lab4"
author: "Sidney Spilfogel"
date: "2024-09-26"
output: "github_document"
---

```{r}
library(car)
library(dplyr)
library(tidyr)
library(ggplot2)
library(psych)
library(bruceR)

#install.packages("car")
```

# Load dataset
```{r}

lab4data <- read.csv("C:/Users/sidne/OneDrive/Documents/GitHub/Lab4/skewed.csv")


```


# Normality
## Normality Plots
```{r}
#how to test for normality

ggplot(lab4data, aes(x = Performance)) + geom_histogram(binwidth = 20) + theme_classic()

ggplot(lab4data, aes(x = Performance)) + geom_density(adjust = 2)  + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance)) + geom_qq()  + theme_classic()

qq+ geom_qq_line()


#functions below are base R functions, they can do the job, but not as customizable as ggplot

hist(lab4data$Performance)

qqnorm(lab4data$Performance, col = "steelblue", lwd = 2)

plot(density(lab4data$Performance, na.rm = TRUE, bw = 90),  lwd=2, main = "")

```

### Normality Plots by Group
```{r}

#a fun function to plot the violin plot by group, but it's not part of ggplot
violinBy(Performance ~ Group, data = lab4data, rain= TRUE, vertical = FALSE) #density plots for each group 


#Find a way to plot the histograms, density, and qq plots by groups using ggplot
ggplot(lab4data, aes(x = Performance)) + geom_histogram(binwidth = 20) + facet_wrap(~Group)+theme_classic()

ggplot(lab4data, aes(x = Performance)) + geom_density(adjust = 2)  +  facet_wrap(~Group) + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance)) + geom_qq()  +  facet_wrap(~Group) + theme_classic()

qq+ geom_qq_line() +  facet_wrap(~Group)

#Use ggplot to create a violin plot by groups


```


## Normality Tests
```{r}

describe(lab4data$Performance)

shapiro.test(lab4data$Performance) #shapiro test tells us data is not normal bc P < 0.05



```

### Normality Tests by Group
```{r}

#Use the describeBy() function to get skewness and kurtosis by group

?describeBy() #shows how to use the function

describeBy(Performance ~ Group, data= lab4data)

#Use the group by function to get shapiro test results by group
lab4data %>%
  group_by(Group) %>% #normality test by group
  summarize(W = shapiro.test(Performance)$statistic, p_value = shapiro.test(Performance)$p.value)

#Use the filter function to get both





```
# Q1: What is your overall conclusion about this variable's normality? Why?

The variables are normal because the p-values are all above 0.05 for each group.It also shows normality because skew and kertosis are within the acceptable range. Despite the graphs not showing normality the tests (shapiro, ketrosis, and skew) show that it is normal.

# Equal Variance between Groups
## Descrptive Variance
```{r}
lab4data_clean <- drop_na(lab4data)

var(lab4data_clean$Performance) #overall variance for group


lab4data_clean %>%
  group_by(Group) %>%
  summarize(variacne = var(Performance)) #variance for each group


```

## Equal Variance Test
```{r}


leveneTest(Performance~Group, lab4data_clean) #can ignore NA data automaticaly so dont need to use clean,levene test is best way - gives varaicne accross all 4 groups

MANOVA(lab4data, dv = "Performance", between = "Group")

#What if you want to test equal variance between 2 groups specifically? 

lab4dataConG1<-lab4data %>% #new data set with filter to give only 2 sets you want to look at 
  filter(Group == "Control" | Group == "G1") #two groups you want to test 

leveneTest(Performance~Group, lab4dataConG1) #run same code to test 


```

# Q2: Overall, does it meet the equal variance assumption across the groups? Why?

No because the variance is not equal across the groups since p < 0.05. Also because some of the groups variance is 10X more than the others. Which is greater than 2x more which is the threshold for equal variance.

# Transformation
```{r}

#if any of the assumption is not met, use transformation 

lab4data$Performance_log <- log10(lab4data$Performance) #adds new column then run t test/anova etc. with that variable 



```

# Q3: Run the above tests again with the transformed outcome. Compare the differences in results. 

The data is normal and variances are equal as shown by the shapiro tests (p>0.05 for each variable) for normality. And the levene test (p>0.05 for each variable) showing equal variance. This shows how transforming the data allowed for the analysis assumptions to be met. 
```{r}

ggplot(lab4data, aes(x = Performance)) + geom_histogram(binwidth = 0.2) + theme_classic()

ggplot(lab4data, aes(x = Performance)) + geom_density(adjust = 2)  + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance_log)) + geom_qq()  + theme_classic()

qq+ geom_qq_line()

hist(lab4data$Performance_log)

qqnorm(lab4data$Performance_log, col = "steelblue", lwd = 2)

plot(density(lab4data$Performance_log, na.rm = TRUE, bw = 90),  lwd=2, main = "")


#a fun function to plot the violin plot by group, but it's not part of ggplot
violinBy(Performance_log ~ Group, data = lab4data, rain= TRUE, vertical = FALSE) #density plots for each group 


#Find a way to plot the histograms, density, and qq plots by groups using ggplot
ggplot(lab4data, aes(x = Performance_log)) + geom_histogram(binwidth = 0.2) + facet_wrap(~Group)+theme_classic()

ggplot(lab4data, aes(x = Performance_log)) + geom_density(adjust = 2)  +  facet_wrap(~Group) + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance_log)) + geom_qq()  +  facet_wrap(~Group) + theme_classic()

qq+ geom_qq_line() +  facet_wrap(~Group)

```

Normality tests
```{r}
#normality tests

describe(lab4data$Performance_log)

shapiro.test(lab4data$Performance_log)

describeBy(Performance_log ~ Group, data= lab4data)

#Use the group by function to get shapiro test results by group
lab4data %>%
  group_by(Group) %>% #normality test by group
  summarize(W = shapiro.test(Performance_log)$statistic, p_value = shapiro.test(Performance_log)$p.value)


```
variance
```{r}

lab4data_clean <- drop_na(lab4data)

var(lab4data_clean$Performance_log) #overall variance for group


lab4data_clean %>%
  group_by(Group) %>%
  summarize(variacne = var(Performance_log)) #variance for each group

leveneTest(Performance_log~Group, lab4data_clean) #can ignore NA data automaticaly so dont need to use clean,levene test is best way - gives varaicne accross all 4 groups

MANOVA(lab4data, dv = "Performance_log", between = "Group")

#What if you want to test equal variance between 2 groups specifically? 

lab4dataConG1<-lab4data %>% #new data set with filter to give only 2 sets you want to look at 
  filter(Group == "Control" | Group == "G1") #two groups you want to test 

leveneTest(Performance~Group, lab4dataConG1) #run same code to test 

```

